### 2023-03-10 intro.qmd
update

The solubility pump is responsible for the great majority of all the anthropogenic $CO_{2}$ that has been absorbed by the ocean and it would be contributing to 10% of the observed DIC gradient according to @Sarmiento2006-ec or 35% according to @Toggweiler2003-ye. Both values are used in the literature (e.g. 10% in @Boyd2019-ez and @Iversen2022-qj and 35% in @Riebesell2009-qj, @Passow2012-zh and @DeVries2022-of).

@@

![Illustration of the three main mechanisms through which the BCP exports carbon to the deep ocean. 1) Production of POC and transport to depth by gravitational settling. 2) Vertical migration of zooplankton that feed at the surface and metabolize the ingested carbon at depth (e.g. production of fecal pellets). 3) Subduction of DOC and suspended POC via mixing processes. Sequestration times for DIC are set depending on the depth at which POC and DOC are exported. Figure taken from @Iversen2022-qj.](image/intro/carbon_pumps_iversen2022.jpeg){#fig-global_carbon_pumps}

@@

Sinking particles are part of a pool of highly diversified materials combining fecal pellets, phytodetritus from sinking phytoplankton, dead zooplankton and organic aggregates referred to as marine snow (i.e. macroscopic organic aggregate \> 0.5 mm) [@Alldredge1988-wm; @Turner2015-af; @Steinberg2017-ri]. Settling velocities for such particles range from 10-100 m day$^{-1}$ up to 1 km day$^{-1}$ for large salp fecal pellets [@Steinberg2017-ri; @Iversen2022-qj].

@@

```{r, Martin Curve}
#| label: fig-Martin-Curve
#| fig-cap: Typical attenuation curve of the POC flux obtained with @eq-Martin-Curve. The color-coded layers represent regions with carbon-specific degradation rate, decreasing with depth. Figure inspired from @Iversen2022-qj.

@@

Based on the Martin curve, @Iversen2022-qj divided the water column in 4 layers (@fig-Martin-Curve): the upper mesopelagic layer (export depth - 300 m) where there is a rapid attenuation of the downward POC flux, the mid-mesopelagic layer (300 - 500 m) and the lower mesopelagic layer (500 - 1,000 m) where both layers show two levels of intermediate attenuation and the bathypelagic layer (1,000 m - seafloor) where the attenuation is relatively uniform. For each layer, @Iversen2022-qj estimated a carbon-specific degradation rate in order to better understand the processes responsible of this attenuation. As a result, @Iversen2022-qj suggests that the POC attenuation in the upper mesopelagic layer is the combined action of microbial degradation and zooplankton that graze on settling aggregates. Here, zooplankton are the main responsible for the attenuation whereas in the middle and lower mesopelagic layers, microbial degradation takes over due to the decrease in zooplankton abundance. Finally, the relatively constant degradation occurring below 1,000 m in the bathypelagic zone could be explained by the increasing pressure and decreasing temperatures on microbial communities.

The impact on the downward POC flux of zooplankton feeding on aggregates are numerous [@Turner2015-af; @Iversen2022-qj]. On the one hand, aggregate feeding, filter-feeding, direct ingestion of fecal pellets (i.e. coprophagy) and fragmentation of large aggregates to small slower-sinking particles will likely decrease the downward POC flux. On the other hand, the ingestion of slow-sinking cells to defecation of large and compact (i.e. fast-settling) fecal pellets will increase it. The remaining organic matter (\~15% according to @Turner2015-af) will then be degraded by microbial communities. In the end, only a small fraction of the exported carbon will reach the sediment where it will be stored for millions of years.

@@

The size of aggregates was usually thought to be a key predictor of settling velocity \[e.g. @Passow2006-hw\] which is coherent with @eq-sinking-speed. However, @Iversen2020-kb did not find evidence for a size-sinking relationship therefore suggesting that excess density would be the main controlling factor. Because the sinking speed determines the time available for degradation in the water column, it also directly impacts the potential for carbon sequestration. With all things considered, there is a thus need for direct measurements of morphological characteristics (size, composition, porosity and density) of particles to better estimate their settling velocities \[@Iversen2020-kb; @Giering2020-zk; @Iversen2022-qj; @Williams2022-xx\].

>>>>

The solubility pump is responsible for the great majority of all the anthropogenic $CO_{2}$ that has been absorbed by the ocean and it would be contributing to 10% of the observed DIC gradient according to @Sarmiento2006-ec or 35% according to @Toggweiler2003-ye. Both values are used in the literature (e.g. 10% in @Boyd2019-ez and @Iversen2023-qj and 35% in @Riebesell2009-qj, @Passow2012-zh and @DeVries2022-of).

@@

![Illustration of the three main mechanisms through which the BCP exports carbon to the deep ocean. 1) Production of POC and transport to depth by gravitational settling. 2) Vertical migration of zooplankton that feed at the surface and metabolize the ingested carbon at depth (e.g. production of fecal pellets). 3) Subduction of DOC and suspended POC via mixing processes. Sequestration times for DIC are set depending on the depth at which POC and DOC are exported. Figure taken from @Iversen2023-qj.](image/intro/carbon_pumps_iversen2022.jpeg){#fig-global_carbon_pumps}

@@

Sinking particles are part of a pool of highly diversified materials combining fecal pellets, phytodetritus from sinking phytoplankton, dead zooplankton and organic aggregates referred to as marine snow (i.e. macroscopic organic aggregate \> 0.5 mm) [@Alldredge1988-wm; @Turner2015-af; @Steinberg2017-ri]. Settling velocities for such particles range from 10-100 m day$^{-1}$ up to 1 km day$^{-1}$ for large salp fecal pellets [@Steinberg2017-ri; @Iversen2023-qj].

@@

```{r, Martin Curve}
#| label: fig-Martin-Curve
#| fig-cap: Typical attenuation curve of the POC flux obtained with @eq-Martin-Curve. The color-coded layers represent regions with carbon-specific degradation rate, decreasing with depth. Figure inspired from @Iversen2023-qj.

@@

Based on the Martin curve, @Iversen2023-qj divided the water column in 4 layers (@fig-Martin-Curve): the upper mesopelagic layer (export depth - 300 m) where there is a rapid attenuation of the downward POC flux, the mid-mesopelagic layer (300 - 500 m) and the lower mesopelagic layer (500 - 1,000 m) where both layers show two levels of intermediate attenuation and the bathypelagic layer (1,000 m - seafloor) where the attenuation is relatively uniform. For each layer, @Iversen2023-qj estimated a carbon-specific degradation rate in order to better understand the processes responsible of this attenuation. As a result, @Iversen2023-qj suggests that the POC attenuation in the upper mesopelagic layer is the combined action of microbial degradation and zooplankton that graze on settling aggregates. Here, zooplankton are the main responsible for the attenuation whereas in the middle and lower mesopelagic layers, microbial degradation takes over due to the decrease in zooplankton abundance. Finally, the relatively constant degradation occurring below 1,000 m in the bathypelagic zone could be explained by the increasing pressure and decreasing temperatures on microbial communities.

The impact on the downward POC flux of zooplankton feeding on aggregates are numerous [@Turner2015-af; @Iversen2023-qj]. On the one hand, aggregate feeding, filter-feeding, direct ingestion of fecal pellets (i.e. coprophagy) and fragmentation of large aggregates to small slower-sinking particles will likely decrease the downward POC flux. On the other hand, the ingestion of slow-sinking cells to defecation of large and compact (i.e. fast-settling) fecal pellets will increase it. The remaining organic matter (\~15% according to @Turner2015-af) will then be degraded by microbial communities. In the end, only a small fraction of the exported carbon will reach the sediment where it will be stored for millions of years.

@@

The size of aggregates was usually thought to be a key predictor of settling velocity [e.g. @Passow2006-hw] which is coherent with @eq-sinking-speed. However, @Iversen2020-kb did not find evidence for a size-sinking relationship therefore suggesting that excess density would be the main controlling factor. Because the sinking speed determines the time available for degradation in the water column, it also directly impacts the potential for carbon sequestration. With all things considered, there is a thus need for direct measurements of morphological characteristics (size, composition, porosity and density) of particles to better estimate their settling velocities [@Iversen2020-kb; @Giering2020-zk; @Iversen2023-qj; @Williams2022-xx].
